Ages of Stars and Planets in the Kepler Field Over the First Four Billion Years

### 2024-09-04 papers/paper/ms.tex
update ms

Cluster membership is a nuanced subject.
Figure~\ref{fig:agescalecompone} is showing reported ages for a set of
spatially and kinematically selected stars that could be cluster
members, or they could be field interlopers.  For NGC6811 and NGC6819,
we adopted candidate members from
\citet{2018A&A...618A..93C,CantatGaudin_2020} and
\citet{Kounkel_2020}.  For Theia~520, we used candidate members from
\citet{Kounkel_2020}.  For Melange-3, we used the candidates reported
by \citet{Barber_2022} and required ``offset'' tangential velocities
below 2\,\kms.  For Cep-Her, we used candidate members from Kerr et
al.~2024 (submitted) with $P_{\rm fin}$$>$0.8.  Even with
contaminants, we can compare the relative age distributions derived by
the different studies because in all instances we are comparing
reported ages against a fixed list of stars.

@@

Figure~\ref{fig:agescalecompone} show that although previous studies
reported ages that agree with the cluster scale for $\gtrsim$1\,Gyr
stars, sub-gigayear stars have historically had their ages
overestimated by 0.3--2\,dex, with severely understated uncertainties.
For isochrone and kinematic ages, this is because these methods rely
on parameters that do not appreciably change at $t$$\lesssim$1\,Gyr.
For the rotational ages in \citet{2023ApJ...952..131M} and
\citet{2024AJ....167..159L}, the discrepancy is caused by details of
the calibration methodologies.  The \texttt{kiauhoku} spin-down model
for instance has known ``stretches'' and ``compressions'' in its age
scale relative to observed open clusters \citep[see][Sec.
7.3]{2023ApJ...952..131M}.  The \citet{Mamajek_2008} calibration used
by \citet{Reinhold_2015} appears more accurate than the former two
models, because it was fitted to reproduce the open cluster sequences
known at the time.  However, the uncertainties from all of these
methods appear to be underestimated.  This is probably because they do
not marginalize over the range of rotation periods accessible to
sub-gigayear stars of a fixed mass and age.

@@

	\vspace{-0.35cm}
    \subfloat{
        \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_preciseagesonly.pdf}
    }
%
%	\vspace{-0.35cm}
%	\subfloat{
%		\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
%	}
  \end{center}
  \vspace{-0.6cm}
  \caption{{\bf Kepler's demographic cliff,} visible in the
  rotation-derived age distributions of its stars (left) and planet
  hosts (middle).  The top row shows \replaced{all stars with temperatures of
  3800--6200\,K for which we calculated rotational ages}{stars with temperatures of 3800--6200\,K}.  Opaque lines
  impose quality cuts on binarity, \added{metallicity, }crowding, and the star's
  evolutionary state; transparent lines do not\added{ (see
  Section~\ref{subsec:flags})}.  The bottom row shows
  stars with temperatures of 4400--5400\,K, which \replaced{have more precise
  ages due to their fast spin-down}{have more precise ages}.  The
  statistical uncertainty for an average star at 1, 2, and 3\,Gyr is
  shown, and is identical across each
  row.  
  Finally, the right panel compares the rotation-derived ages
  against star formation histories derived using CMD fitting (dashed
  and dotted lines).  Completeness in Kepler's $P$$_{\rm rot}$
  detection sensitivity is near unity until $t$$\lesssim$3\,Gyr\added{, and decreases rapidly at older ages}
  \citep{2022ApJ...937...94M}.
  \label{fig:hist_tgyro}
  }
\end{figure*}

>>>>

Cluster membership is a nuanced subject.
Figure~\ref{fig:agescalecompone} is showing reported ages for a set of
spatially and kinematically selected stars that could be cluster
members, or they could be field interlopers.  For NGC6811 and NGC6819,
we adopted candidate members from
\citet{2018A&A...618A..93C,CantatGaudin_2020} and
\citet{Kounkel_2020}.  For Theia~520, we used candidate members from
\citet{Kounkel_2020}.  For Melange-3, we used the candidates reported
by \citet{Barber_2022} and required ``offset'' tangential velocities
below 2\,\kms.  For Cep-Her, we used candidate members from
\citet{Kerr2024} with $P_{\rm fin}$$>$0.8.  Even with
contaminants, we can compare the relative age distributions derived by
the different studies because in all instances we are comparing
reported ages against a fixed list of stars.

@@

Figure~\ref{fig:agescalecompone} show that although previous studies
reported ages that agree with the cluster scale for $\gtrsim$1\,Gyr
stars, sub-gigayear stars have historically had their ages
overestimated by 0.3--2\,dex, with severely understated uncertainties.
For isochrone and kinematic ages, this is because these methods rely
on parameters that do not appreciably change at $t$$\lesssim$1\,Gyr.
For the rotational ages in \citet{2023ApJ...952..131M} and
\citet{2024AJ....167..159L}, the discrepancy is caused by details of
the calibration methodologies.  The \texttt{kiauhoku} spin-down model
for instance has known ``stretches'' and ``compressions'' in its age
scale relative to observed open clusters \citep[see][Sec.~7.3]{2023ApJ...952..131M}.  The \citet{Mamajek_2008} calibration used
by \citet{Reinhold_2015} appears more accurate than the former two
models, because it was fitted to reproduce the open cluster sequences
known at the time.  However, the uncertainties from all of these
methods appear to be underestimated.  This is probably because they do
not marginalize over the range of rotation periods accessible to
sub-gigayear stars of a fixed mass and age.

@@

	\vspace{-0.35cm}
    \subfloat{
        \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_preciseagesonly.pdf}
    }
%
%	\vspace{-0.35cm}
%	\subfloat{
%		\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
%	}
  \end{center}
  \vspace{-0.6cm}
  \caption{{\bf Kepler's demographic cliff,} visible in the
  rotation-derived age distributions of its stars (left) and planet
  hosts (middle).  The top row shows \replaced{all stars with temperatures of
  3800--6200\,K for which we calculated rotational ages}{stars with temperatures of 3800--6200\,K}.  Opaque lines
  impose quality cuts on binarity, \added{metallicity, }crowding, and the star's
  evolutionary state; transparent lines do not\added{ (see
  Section~\ref{subsec:flags})}.  The bottom row shows
  stars with temperatures of 4400--5400\,K, which \replaced{have more precise
  ages due to their fast spin-down}{have more precise ages}.  The
  statistical uncertainty for an average star at 1, 2, and 3\,Gyr is
  shown, and is identical across each
  row.  
  Finally, the right panel compares the \replaced{rotation-derived}{rotational} ages
  against star formation histories derived \replaced{using CMD
  fitting}{from CMD fits to other regions of the Galaxy} (dashed
  and dotted lines).  Completeness in Kepler's $P$$_{\rm rot}$
  detection sensitivity is near unity until $t$$\lesssim$3\,Gyr\added{, and decreases rapidly at older ages}
  \citep{2022ApJ...937...94M}.
  \label{fig:hist_tgyro}
  }
\end{figure*}
